
\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{\texorpdfstring{\color{gray}Acknowledgements}{Acknowledgements}}
%\addchap{\texorpdfstring{\color{gray}Acknowledgments}{Acknowledgments2}}
%\addchap{\hyperref[sec:toc]{\color{gray}Acknowledgments}}

% personal acknowledgements
There are some people I want to thank, without whom the realization and results of my thesis certainly would have been different.
% Volker
First and foremost, I thank my PhD supervisor Volker~Bothmer for supervising my doctorate, for letting me work so freely, and for providing helpful discussions and comments.
% thesis committee
I am grateful to Volker~Bothmer and Ansgar~Reiners for constituting my advisory committee and also for kindly agreeing to evaluate my work as thesis committee.
% further committee members
% I thank the further committee members Andreas~Tilgner, XXX, XXX, and XXX who kindly agreed to evaluate my thesis defense.

% IAG
My appreciation goes to all members of the Institute for Astrophysics for facilitating such a pleasant work environment.
I thank all current and former associates of the space weather group for their friendly company on my way -- in particular Jens, Johannes, Niclas, and Giuseppe for our lively discussions, which were every now and then even scientifically motivated ;)

% my office mates
I especially thank my eternal office mates Eckhard, Jonas, and Adam:
Eckhard for instantly integrating me at the start of my PhD and the good friendship since,
Jonas for the interesting debates during the beginning,
and Adam for accompanying me through the last difficulties on the way.

I acknowledge the office for cozily accomodating me all the long hours, especially the couch for giving me strength, and the Shadow Worker for guarding the office. I honor the brands Star~Wars and LEGO for providing lots of merchandise and distractions.

% family and friends
I am grateful to my family and friends for supporting me at all times.
% else
Those who are not mentioned above but deserve it are thanked hereby as well.
% Sun
A shout-out to the Sun for creating the solar wind and harboring all life we know of (yet).\\

(I am thankful to the proofreaders for reading drafts of this manuscript and providing very helpful comments and suggestions. Volker, Adam, and Tal?)\\


-------------\\
% academic acknowledgments

for creating and making scientific data available:\\
- SWS list (see Acknowledgements in Elliott2013; directly Ian~Richardson)\\
- OMNI data (see paper)\\
- ACE data\\
- Helios data (see paper)\\
- Kp data\\
- SSN data\\

I acknowledge financial support from the following projects:\\
AFFECTS: Solar wind correlation with magnetosphere\\
CGAUSS: Solar wind model and extrapolation\\
HELCATS: Minimum variance analyses of magnetic clouds\\
OPTIMAP: Solar wind ACE time series\\


This work was done within the scope of the CGAUSS project, funded by the German Aerospace Center (grant number...).\\
The authors thank the Helios and OMNI teams for creating and making availabe the solar wind in situ data. The Helios and the OMNI data sets were supplied by the Space Physics Data Facility at NASA's Goddard Space Flight Center.\\
We thank the 'referees' for /important comments and suggestions. /careful review of this paper.\\

facilities/persons which made available the Helios and OMNI data\\
``Ian Richardson and Hilary Cane for making their Interplanetary Coronal Mass Ejection list readily available''\\
``Joseph King and Natalia Papitashvili for creating the combined OMNI data set''\\
The OMNI data were downloaded from NASA's Space Physics Data Facility ...\\
``We thank the reviewers for the careful review of this paper''\\
The Authors extend their thanks to the referee for important comments and suggestions.\\
This work was supported by the DLR CGAUSS project\\
This work has been done in the frame of the CGAUSS project (url), funded by the DLR\\
We would like to thank the Helios team and NASA's SPDF for supplying the plasma and magnetic field data (url)\\


The results presented in this thesis rely on the \Kp{}~index, calculated and made available by the German Research Centre for Geosciences in Potsdam from data collected at magnetic observatories. We thank the involved national institutes, the INTERMAGNET network and ISGI (isgi.unistra.fr).\\



acknowledgements to MAG/SWEPAM Team for ACE data in basics plots...\\

Helios data acknowledgments:\\
Solar wind data courtesy of R.~Schwenn, Max-Planck-Institut für Aeronomie, Lindau, magnetic field data courtesy of F.~Neubauer, Universität zu Köln. (see paper...)\\

Data acknowledgments for \autoref{chap:chapter2}:\\
%\section{Acknowledgments}
The research leading to these results has received funding from the European Union's Seventh Framework Programme (FP7/2007-2013) under the grant agreement number 263506 (AFFECTS). The results presented in this paper rely on the \Kp{}~index, calculated and made available by the German Research Centre for Geosciences in Potsdam from data collected at magnetic observatories. We thank the involved national institutes, the INTERMAGNET network and ISGI (isgi.unistra.fr). The authors thank the OMNI PIs/teams for creating and making available the solar wind in-situ data. The OMNI data are supplied by the NASA Space Science Data Coordinated Archive and the Space Physics Data Facility at NASA's Goddard Space Flight Center. Additional thanks for maintaining and providing the international sunspot number series goes to the World Data Center -- Sunspot Index and Long-term Solar Observations at the Solar Influences Data Analysis Center (SIDC), Royal Observatory of Belgium. The update of the hourly solar wind structure list was kindly provided by Ian~Richardson of the NASA Goddard Space Flight Center and CRESST/University of Maryland via the CEDAR Database at the National Center for Atmospheric Research, which is supported by the National Science Foundation.\\

% NASA ADS
This research has made use of NASA's Astrophysics Data System Bibliographic Services.\\

