
\chapter*{Curriculum vitae}
\addcontentsline{toc}{chapter}{\texorpdfstring{\color{gray}Curriculum vitae}{Curriculum vitae}}
%\addchap{\texorpdfstring{\color{gray}Curriculum vitae}{Curriculum vitae2}}

%compulsory for printed version
%can be omitted for the eDiss version

Malte~Venzmer, born 5~February 1984 in Bremerhaven, Lower Saxony, Germany, finished secondary school (Abitur) at the Gymnasium Ganderkesee in 2003. Subsequently he performed his civilian service working in a care home for disabled persons in Westerland, Sylt.

In 2004 he began to study physics at the University of Konstanz. During his studies, in summer 2008, he did an internship at the \AA{}ngström Laboratory, Uppsala University, Sweden, working in the field of materials physics. In 2010/2011 he wrote his diploma thesis in the \textit{Radio- and X-ray astronomy} group of Dr.~Jürgen~Kerp at the Argelander-Institute for Astronomy, University of Bonn. There, following his diploma in physics, he continued working under a scholarship until end of 2011.

In 2012 he started as a PhD candidate in the \textit{Solar, heliospheric and space weather research} group of Dr.~Volker~Bothmer at the Institute for Astrophysics, University of Göttingen. During his doctoral studies, he worked as a research assistent for several national and international projects. Most of his main results are presented in this very thesis.

%moke setup, electrical steels
%radio astronomy, molecular hydrogen clouds, interaction of Magellanic System with the Milky Way
%solar wind, geomagnetic storms


% \vspace{\baselineskip}
% 
% \begin{large}Malte~Venzmer
% \end{large}
% \begin{description}[itemsep=2ex, leftmargin=2.7cm, style=sameline]
% 	\item[5~February 1984] born in Bremerhaven, Lower Saxony, Germany
% 	\item[2003] finished secondary school (Abitur) at the Gymnasium Ganderkesee
% 	\item[2004] began to study physics at the University of Konstanz
% 	\item[Summer 2008] internship at the \AA{}ngström Laboratory, Uppsala University, Sweden, working in the field of materials physics
% 	\item[2010/2011] wrote diploma thesis in the \textit{Radio- and X-ray astronomy} group of Dr.~Jürgen~Kerp at the Argelander-Institute for Astronomy, University of Bonn
% 	\item[Until end of 2011] continued working under a scholarship
% 	\item[2012] started as a PhD candidate in the \textit{Solar, heliospheric and space weather research} group of Dr.~Volker~Bothmer at the Institute for Astrophysics, University of Göttingen
% 	\item[During doctoral studies] worked as a research assistent for several national and international projects
% \end{description}
